---
title: "Accuracy of DeepLoc for different score cutoffs."
output: html_notebook
---

```{r}
library(tidyverse)
library(assertthat)
library(parallel)
```

```{r}
location_mapping = read_tsv("../results/location_mapping.tsv")
hpa = read_tsv("../results/hpa_filtered.tsv")
deeploc = read_tsv("../results/deeploc_all.tsv")
deeploc_fil = read_tsv("../results/deeploc_filtered.tsv")
```

```{r}
hpa_mapped = hpa %>%
  inner_join(location_mapping, by=c(subcellular_location="target_location")) %>%
  filter(target_source == 'hpa')

deeploc_clean = deeploc %>%
  inner_join(deeploc_fil) %>%
  select(-`#ID`, -uniprot, -`Membrane-bound`, -isoform) %>%
  rowwise() %>% 
  mutate(max_score=max(Nucleus:Peroxisome)) %>%
  distinct(hgnc, .keep_all=TRUE) 

deeploc_val = deeploc_clean %>% 
  inner_join(hpa_mapped, by=c(hgnc="hgnc")) %>%
  rowwise() %>%
  mutate(true_positive=Prediction==deeploc_location) %>%
  ungroup() %>%
  group_by(hgnc, max_score) %>%
  summarise(true_positive=any(true_positive))
```

Show a plot similar to the one from the LocTree2 paper: Accuracy depending on the cutoff. 

Dataset: The 7821 proteins which are both in HPA and deeploc and had a location that could be mappeed. 
Coverage is defined as the fraction of predictions that exceed a certain score. 
Accuracy is defined as the number of predictions that are correct divided by the total number of predictions.
A prediction is correct if the predicted class matches to one of the annotated classes of the HPA. 
```{r}
cutoffs = seq(0, .99, 0.01)

by_cutoff = function(thres) {
  above_thres = deeploc_val %>% filter(max_score >= thres)
  tp = above_thres %>% filter(true_positive)
  
  acc = nrow(tp) / nrow(above_thres)
    
  coverage = nrow(above_thres) / nrow(deeploc_val)
  
  return(c(accuracy=acc, coverage=coverage))
}

res = mcmapply(by_cutoff, cutoffs, mc.cores=4)

res_df = tbl_df(t(res)) %>%
  mutate(cutoff=cutoffs) %>% 
  gather(type, score, -cutoff)

ggplot(res_df, aes(x=cutoff, y=score)) + geom_line(aes(colour=type))
```