---
title: "DeepLoc Multiclass predictions - Part II"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggpubr)
library(cowplot)
library(pROC)
```

Before, we have shown that if we *know*, how many locations a protein has, we are 
able to predict these locations from the DeepLoc scores at ~50% accuracy. 

In this notebook, we will investigate, if we are also capable of *predicting* at how
many locations a protein occurs. 

# Load and prepare data
We use the HPA as a reference dataset. 
We map the deeploc data to the HPA data.
```{r, warning=FALSE, message=FALSE}
deeploc = read_tsv("../results/deeploc_all.tsv")
location_mapping = read_tsv("../results/location_mapping.tsv")
hpa = read_tsv("../results/hpa_filtered.tsv")
```

```{r, include=FALSE}
# subcellular locations both in deeploc and hpa (mapped to deeploc)
shared_locations = location_mapping %>%
  filter(target_source == 'hpa') %>%
  select(deeploc_location) %>%
  distinct()

# protein identifiers both in deeploc and hpa
shared_prots = deeploc %>% 
  inner_join(shared_locations, by=c(Prediction="deeploc_location")) %>%
  inner_join(hpa, by=c(hgnc='hgnc')) %>%
  select(hgnc) %>%
  distinct()

nrow(shared_prots)
paste(shared_locations$deeploc_location, sep=",", collapse = ", ")
```

```{r, include=FALSE}
# hpa mapped to deeploc locations, filtered for proteins also in deeploc dataset
hpa_fil = hpa %>%
  inner_join(
    filter(location_mapping, target_source == 'hpa'),
    by=c(subcellular_location="target_location")) %>%
  inner_join(shared_prots) %>%
  inner_join(shared_locations) %>%
  select(hgnc, deeploc_location) %>%
  distinct()

deeploc_fil = deeploc %>%
  inner_join(shared_prots) %>%
  inner_join(shared_locations, by=c(Prediction="deeploc_location")) %>%
  select(hgnc, deeploc_location="Prediction") %>%
  distinct()
```

```{r, include=FALSE}
hpa_loc_count = hpa_fil %>%
  group_by(hgnc) %>%
  summarise(n_locations=n()) %>%
  select(hgnc, n_locations)

deeploc_scores = deeploc %>% 
  filter(isoform == 1 | is.na(isoform)) %>% # only canonical isoforms
  inner_join(hpa_loc_count) %>%
  select(-`#ID`, -isoform, -`Membrane-bound`, -uniprot) %>%
  gather(compartment, score, -hgnc, -Prediction, -n_locations) %>%
  arrange(hgnc) %>%
  distinct(hgnc, compartment, .keep_all=TRUE)
```
After preparation we have a table like this:
```{r}
deeploc_scores
```

# Results
## Distribution of Scores
Here, we show the score distribution for the highest, second highest, ... prediction
for proteins that have one or two locations according to HPA respectively. 

While there are (statistically significant) differences, we already see 
that it will be hard to distinguish between the two categories using the 
score only. 
```{r}
deeploc_score_distrib = deeploc_scores %>%
  filter(n_locations <= 2) %>%
  mutate(n_locations=as.factor(n_locations)) %>%
  group_by(hgnc) %>%
  mutate(rank=rank(-score, ties.method="max")) %>%
  arrange(hgnc)

deeploc_score_distrib %>%
  ggplot(aes(x=as.factor(rank), y=score)) + geom_boxplot(aes(colour=n_locations))
```

## Simple classification by Threshold
The basic idea is, to use a threshold for the highest score. 
Above a certain threshold, we consider the protein as "having only one subcellular location". 
Below the threshold, we consider it as "having two subcellular locations". 

The figur below shows an example for a cutoff of 0.7:
```{r}
thres = .7

get_state = function(n_pred, n_actual) {
  if(n_pred == 1) {
    if(n_actual == 1) {
      return("TN")
    } else {
      return("FN")
    }
  } else {
    if(n_actual == 2) {
      return("TP")
    } else{
      return("FP")
    }
  }
}

max_score = deeploc_score_distrib %>%
  filter(rank == 1) %>% 
  mutate(pred_locations=ifelse(score>thres,1,2)) %>%
  mutate(state=get_state(pred_locations, n_locations))

max_score %>%
  ggplot(aes(x=n_locations, y=score)) +
  geom_violin(col='black') +
  geom_boxplot(width=.1) + 
  geom_jitter(alpha=.3, shape=1, aes(colour=state)) +
  geom_hline(yintercept=thres, col='red')
```

## ROC Analysis
To quantify the performance for all thresholds, we perform a ROC analysis. 
As expected, the overall performance is quite poor with an AUC of 0.54. 
```{r}
roc_curve = roc(response=max_score$n_locations, predictor=max_score$score)
plot(roc_curve)
title(paste("AUC =", round(roc_curve$auc, 2)))
```

# Conclusion
We conclude that multiclass prediction from DeepLoc scores alone is not feasible. 
For further investigation we suggest either

 * obtaining unnormalized scores from the Neural Network or
 * adding the number of locations as an output to the Neural Network directly. 
 
