---
title: "DeepLoc Multiclass predictions"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(ggpubr)
library(cowplot)
```

In this notebook, we try to investigate how DeepLoc handles proteins, that occur at multiple locations. 

### Load dataset
We use the HPA as a reference dataset
```{r}
deeploc = read_tsv("../results/deeploc_all.tsv")
location_mapping = read_tsv("../results/location_mapping.tsv")
hpa = read_tsv("../results/hpa_filtered.tsv")
```

Do the location mapping and some other preparatory work:
```{r}
# subcellular locations both in deeploc and hpa (mapped to deeploc)
shared_locations = location_mapping %>%
  filter(target_source == 'hpa') %>%
  select(deeploc_location) %>%
  distinct()

# protein identifiers both in deeploc and hpa
shared_prots = deeploc %>% 
  inner_join(shared_locations, by=c(Prediction="deeploc_location")) %>%
  inner_join(hpa, by=c(hgnc='hgnc')) %>%
  select(hgnc) %>%
  distinct()

nrow(shared_prots)
paste(shared_locations$deeploc_location, sep=",", collapse = ", ")
```

```{r}
# hpa mapped to deeploc locations, filtered for proteins also in deeploc dataset
hpa_fil = hpa %>%
  inner_join(
    filter(location_mapping, target_source == 'hpa'),
    by=c(subcellular_location="target_location")) %>%
  inner_join(shared_prots) %>%
  inner_join(shared_locations) %>%
  select(hgnc, deeploc_location) %>%
  distinct()

deeploc_fil = deeploc %>%
  inner_join(shared_prots) %>%
  inner_join(shared_locations, by=c(Prediction="deeploc_location")) %>%
  select(hgnc, deeploc_location="Prediction") %>%
  distinct()
```

# Find proteins with multiple isoforms
We extract all proteins with multiple isoforms from the deeploc dataset and look if the multiple locations somehow correspond to the corresponding multiclass predictions. 

```{r}
deeploc_multi_prots = deeploc_fil %>%
  group_by(hgnc) %>%
  count() %>%
  filter(n >= 2) %>%
  select(hgnc)

deeploc_validation = deeploc_fil %>%
  inner_join(deeploc_multi_prots) %>%
  left_join(mutate(hpa_fil, in_hpa=1),
            by=c(hgnc="hgnc", deeploc_location="deeploc_location")) %>% 
  arrange(hgnc, deeploc_location)

deeploc_acc = deeploc_validation %>%
  group_by(hgnc) %>%
  summarise(true=sum(in_hpa, na.rm=TRUE), actual=n())

true_positive = deeploc_acc %>%
  filter(true==actual) %>%
  nrow()

true_positive / nrow(deeploc_multi_prots)
```
random with 8 categories: 12.5 (approx.). No statistical test, but low effect...

```{r}
hpa_multi_prots = hpa_fil %>%
  group_by(hgnc) %>%
  count() %>%
  filter(n >= 2) %>%
  select(hgnc)

hpa_validation = hpa_fil %>%
  inner_join(hpa_multi_prots) %>%
  left_join(mutate(deeploc_fil, in_deeploc=1),
            by=c(hgnc="hgnc", deeploc_location="deeploc_location")) %>% 
  arrange(hgnc, deeploc_location)

hpa_acc = hpa_validation %>%
  group_by(hgnc) %>%
  summarise(true=sum(in_deeploc, na.rm=TRUE), actual=n())

true_positive = hpa_acc %>%
  filter(true==actual) %>%
  nrow()

true_positive / nrow(hpa_multi_prots)
```

Even less, because many multi-loc proteins proably don't have any isoforms. 


# Find proteins with multiple locations (hpa)
To it the other way round... find all proteins that are annotated as multi-loc. We look at the predictions of DeepLoc. How do the scores look like? 

e.g. max score multi vs. max-score all


```{r}
hpa_loc_count = hpa_fil %>%
  group_by(hgnc) %>%
  summarise(n_locations=n()) %>%
  select(hgnc, n_locations)

deeploc_scores = deeploc %>% 
  filter(isoform == 1 | is.na(isoform)) %>% # only canonical isoforms
  inner_join(hpa_loc_count) %>%
  select(-`#ID`, -isoform, -`Membrane-bound`, -uniprot) %>%
  gather(compartment, score, -hgnc, -Prediction, -n_locations) %>%
  arrange(hgnc) %>%
  distinct(hgnc, compartment, .keep_all=TRUE)
```

```{r}
max_score = deeploc_scores %>%
  mutate(n_locations=as.factor(n_locations)) %>%
  group_by(hgnc, n_locations) %>%
  summarise(score=max(score)) 

max_score %>%
  ggplot(aes(x=n_locations, y=score)) +
  geom_boxplot() + 
  stat_compare_means(comparisons = list(c(1,2), c(1,3), c(1,4)), method = 'wilcox.test') + 
  ggtitle("Maximum score")
```


```{r}
max_score = deeploc_scores %>%
  mutate(n_locations=as.factor(n_locations)) %>%
  group_by(hgnc, n_locations) %>%
  summarise(score=sort(score, decreasing=TRUE)[2])

max_score %>%
  ggplot(aes(x=n_locations, y=score)) + 
  geom_boxplot() + 
  stat_compare_means(comparisons = list(c(1,2), c(1,3), c(1,4)), method = 'wilcox.test') + 
  ggtitle("rank 2 scores")
```

As a simple test, let's see if including the second-highest prediction for multi-class proteins yields something reasonable: 
```{r}
deeploc_multi_pred = deeploc_scores %>%
  filter(n_locations >= 2) %>%
  group_by(hgnc) %>%
  mutate(rank=rank(-score)) %>%
  filter(rank <= 2) %>%
  arrange(hgnc)

deeploc_multi_validation = deeploc_multi_pred %>%
  left_join(mutate(hpa_fil, in_hpa=1),
            by=c(hgnc="hgnc", compartment="deeploc_location")) %>% 
  arrange(hgnc, -score)

deeploc_acc = deeploc_multi_validation %>%
  group_by(hgnc) %>%
  summarise(true=sum(in_hpa, na.rm=TRUE), actual=n())

true_positive = deeploc_acc %>%
  filter(true==actual) %>%
  nrow()

true_positive / nrow(deeploc_acc)
```


